---
title: "Foxa2_tracing - quantifications and plotting"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r message=FALSE}
# Load all packages
rm(list=ls())

setwd("~/dev/repos/tdTom_LinTracing/R_analysis/")

library(tidyverse)
library(readxl)
library(pheatmap)
library(RColorBrewer)
library(lemon)
library(viridis)

```

### Load data

```{r}
clean_segment <- read.csv("LinTraced_segment_plot.csv")
  
```

### Find thresholds
```{r fig.width=5, fig.height=4}

sox2_thres = 90
nkx22_thress = 90
tdt_thress = 50

ggplot(clean_segment, aes(x=SOX2_avg_signal, y=NKX22_ave_signal)) +
  geom_hex(binwidth = c(0.04, 0.04))  +
  geom_hline(yintercept=nkx22_thress,  linetype="dashed", color = "red") +
  geom_vline(xintercept=sox2_thres,  linetype="dashed", color = "red") +
  scale_x_log10()+
  scale_y_log10() +
  facet_wrap(~ embryoID) +
  theme_bw(base_size = 12)

ggplot(clean_segment, aes(x=SOX2_avg_signal, y=NKX22_ave_signal)) +
  geom_point(size=0.2, alpha=0.5) +
  geom_hline(yintercept=nkx22_thress,  linetype="dashed", color = "red") +
  geom_vline(xintercept=sox2_thres,  linetype="dashed", color = "red") +
  scale_x_log10()+
  scale_y_log10() +
  facet_wrap(~ embryoID) +
  theme_bw(base_size = 12)

```

#### tdTomato

```{r}

ggplot(clean_segment, aes(x=tdTom_ave_signal, y=NKX22_ave_signal)) +
  geom_point(size=0.2, aes(color=log10(tdTom_ave_signal))) +
  scale_color_gradient2(low = "grey", mid = "white", high = "brown") +
  geom_hline(yintercept=nkx22_thress,  linetype="dashed", color = "blue") +
  geom_vline(xintercept=tdt_thress,  linetype="dashed", color = "blue") +
  scale_x_log10()+
  scale_y_log10() +
  facet_wrap(~ embryoID) +
  theme_bw(base_size = 12)

ggplot(clean_segment, aes(x=tdTom_ave_signal, y=SOX2_avg_signal)) +
  geom_point(size=0.2, aes(color=log10(tdTom_ave_signal))) +
  scale_color_gradient2(low = "grey", mid = "white", high = "brown") +
  geom_hline(yintercept=sox2_thres,  linetype="dashed", color = "blue") +
  geom_vline(xintercept=tdt_thress,  linetype="dashed", color = "blue") +
  scale_x_log10()+
  scale_y_log10() +
  facet_wrap(~ embryoID) +
  theme_bw(base_size = 12)

```

### Plot cell type based on markers

Use the position as a QC for how well the thresholds identify cell types.
Nkx2.2 has background signal in endothelial cells because it's a mouse antibody. Filter by positions in the next step. 

```{r}
full_classified <- clean_segment %>%
  mutate(celltype=if_else(SOX2_avg_signal > sox2_thres & NKX22_ave_signal > nkx22_thress, "p3",
                          if_else(NKX22_ave_signal > nkx22_thress, "V3",
                                  if_else(SOX2_avg_signal > sox2_thres,"Sox2_notp3","othercell")))) %>%
         filter(elipse_x <1000, elipse_y<1500, elipse_x>0,elipse_y >0)


ggplot(full_classified, aes(x=elipse_x, y=-elipse_y)) +
  geom_point(size=0.2,aes(color=celltype)) +
  scale_color_manual(values = c("Lightgrey","Darkgreen","DarkGrey","Magenta")) +
  facet_wrap(~ embryoID) +
  coord_fixed(ratio = 1) +
  theme_bw(base_size = 8)

```


### Filter background Nkx22 cells based on position

Nkx2.2-expressing cells that are positioned more than 2 standard deviations away from the mean (dashed red line) will be excluded. Visualize the excluded cells below.

```{r}

full_classified_filter_nkx22 <- full_classified %>%
  filter(celltype %in% c("V3","p3")) %>%
  group_by(Image) %>%
  mutate(ave_nkx22_y = mean(elipse_y, na.rm = TRUE),
         sd_nkx22_y = sd(elipse_y, na.rm = TRUE),
         dev_y = abs(elipse_y-ave_nkx22_y))

dotsize=0.2


ggplot(full_classified_filter_nkx22, aes(x=elipse_x, y=-elipse_y)) +
  geom_point(size=dotsize,aes(color=NKX22_ave_signal)) +
  scale_color_continuous(type = "viridis") +
  expand_limits(y=c(0,-1500),x=c(0,1000)) +
  geom_hline(aes(yintercept=-(ave_nkx22_y+2*sd_nkx22_y)),linetype='dashed', col = 'red') +
  facet_wrap(~ embryoID) +
  coord_fixed(ratio = 1) +
  theme_bw(base_size = 8)

```

Position filter for Nkx22 cells per Image

```{r}
nkx22_position_filter <- full_classified_filter_nkx22 %>%
  mutate(nkx22_y2sd_filter=ave_nkx22_y+2*sd_nkx22_y) %>% 
  select(Image,embryoID,nkx22_y2sd_filter) %>% unique()
  
```



Re-classify adding the positional filter

```{r}

classified_filtered <- clean_segment %>%
  left_join(nkx22_position_filter, by = c("Image","embryoID")) %>% 
  mutate(celltype=if_else(SOX2_avg_signal > sox2_thres & NKX22_ave_signal > nkx22_thress & elipse_y < nkx22_y2sd_filter, "p3",
                          if_else(NKX22_ave_signal > nkx22_thress & elipse_y < nkx22_y2sd_filter, "V3",
                                  if_else(SOX2_avg_signal > sox2_thres,"Sox2_notp3","othercell")))) %>%
         filter(elipse_x <1000, elipse_y<1500, elipse_x>0,elipse_y >0)


ggplot(classified_filtered, aes(x=elipse_x, y=-elipse_y)) +
  geom_point(size=0.2,aes(color=celltype)) +
  scale_color_manual(values = c("Lightgrey","Darkgreen","DarkGrey","Magenta")) +
  facet_wrap(~ embryoID) +
  coord_fixed(ratio = 1) +
  theme_bw(base_size = 8)

```


### Classify traced cells 

```{r}
full_classified_tdtom  <- classified_filtered %>%
           group_by(Image) %>%
           mutate(tdtom_status = if_else(tdTom_ave_signal > tdt_thress,"traced","untraced")) 

```

QC plot "traced". 

We see that they are in the expected positions.

```{r}
ggplot(full_classified_tdtom, aes(x=elipse_x, y=-elipse_y)) +
  geom_point(size=0.2,aes(color=tdtom_status)) +
  facet_wrap(~ embryoID) +
  coord_fixed(ratio = 1) +
  theme_bw(base_size = 8)
```

QC plot "traced" only by cell type
```{r}
ggplot(full_classified_tdtom %>% filter(tdtom_status=="traced"), aes(x=elipse_x, y=-elipse_y)) +
  geom_point(size=0.2,aes(color=celltype), alpha=0.5) +
  scale_color_manual(values = c("Lightgrey","Darkgreen","orange","Magenta")) +
  facet_wrap(~ embryoID) +
  coord_fixed(ratio = 1) +
  theme_bw(base_size = 8)

```

## Count traced cells

How many cells in each section is labelled as p3 or V3 and is tdTom traced?

Use `count` to plot zero counts.

```{r fig.width=5, fig.height=3}
full_classified_tdtom_counts <- full_classified_tdtom %>%
  filter(tdtom_status=="traced") %>%
  mutate(celltype = factor(celltype, levels = c("p3","V3","Sox2_notp3","othercell"))) %>%
  group_by(embryoID, position, tam) %>%
  count(celltype,.drop=FALSE)


ggplot(full_classified_tdtom_counts %>% filter(celltype %in% c("p3","V3")), aes(x=tam, y=n)) +
  geom_col(aes(fill=embryoID),position=position_dodge2(0.9, preserve = "single")) +
  geom_point(aes(fill=embryoID), color="black", shape=21,position=position_dodge2(0.9, preserve = "single")) +
  facet_wrap(~ celltype) +
  ylab("Number of tdTomato+ cells per section") +
  xlab("Time of tamoxifen administration") +
  theme_bw(base_size = 12)

```



```{r}
sessionInfo()
```

